---
title: "trollExp - Run a TROLL experiment with soil"
author: Sylvain Schmitt
date: May 16, 2023
output:
  github_document: 
    toc: yes
    toc_depth: 2
---

```{r setup, include=FALSE}
rm(list = ls()) ; invisible(gc()) ; set.seed(42)
library(knitr)
library(tidyverse)
theme_set(bayesplot::theme_default())
opts_chunk$set(echo = F, message = F, warning = F, fig.height = 6, fig.width = 8,
               cache = F, cache.lazy = F)
```

[`singularity` & `snakemake`](https://github.com/sylvainschmitt/snakemake_singularity) workflow to run a TROLL experiment with soil.

```{r dag, fig.cap="Workflow."}
knitr::include_graphics("dag/dag.svg")
```

# Installation

- [x] Python ≥3.5
- [x] Snakemake ≥5.24.1
- [x] Golang ≥1.15.2
- [x] Singularity ≥3.7.3
- [x] This workflow

```{bash, eval=F, echo=T}
# Python
sudo apt-get install python3.5
# Snakemake
sudo apt install snakemake`
# Golang
export VERSION=1.15.8 OS=linux ARCH=amd64  # change this as you need
wget -O /tmp/go${VERSION}.${OS}-${ARCH}.tar.gz https://dl.google.com/go/go${VERSION}.${OS}-${ARCH}.tar.gz && \
sudo tar -C /usr/local -xzf /tmp/go${VERSION}.${OS}-${ARCH}.tar.gz
echo 'export GOPATH=${HOME}/go' >> ~/.bashrc && \
echo 'export PATH=/usr/local/go/bin:${PATH}:${GOPATH}/bin' >> ~/.bashrc && \
source ~/.bashrc
# Singularity
mkdir -p ${GOPATH}/src/github.com/sylabs && \
  cd ${GOPATH}/src/github.com/sylabs && \
  git clone https://github.com/sylabs/singularity.git && \
  cd singularity
git checkout v3.7.3
cd ${GOPATH}/src/github.com/sylabs/singularity && \
  ./mconfig && \
  cd ./builddir && \
  make && \
  sudo make install
# detect Mutations
git clone git@github.com:sylvainschmitt/trollExp.git
cd trollExp
git checkout soil
```

# Usage

## Locally

```{bash, eval=F, echo=T}
snakemake -np -j 1 # dry run
snakemake --dag | dot -Tsvg > dag/dag.svg # dag
snakemake -j 1 --use-singularity --singularity-args "\-e" # run
```

## HPC

### Muse

```{bash, eval=F, echo=T}
module load snakemake # for test on node
snakemake -np # dry run
sbatch job_muse.sh # run
```

### Genologin

```{bash, eval=F, echo=T}
module load bioinfo/snakemake-5.25.0 # for test on node
snakemake -np # dry run
sbatch job_genologin.sh # run
```

# Workflow

## Soil

### [get_soil](https://github.com/sylvainschmitt/trollExp/blob/soil/rules/get_soil.smk)

* Script: [`get_soil.R`](https://github.com/sylvainschmitt/trollExp/blob/soil/scripts/get_soilx.R)

Download SoildGrids data for a given variable and depth.

### [extract_soil](https://github.com/sylvainschmitt/trollExp/blob/soil/rules/extract_soil.smk)

* Script: [`extract_soil.R`](https://github.com/sylvainschmitt/trollExp/blob/soil/scripts/extract_soil.R)

Extract soil data for given addresses.

## Run TROLL

### [troll_warm](https://github.com/sylvainschmitt/trollExp/blob/soil/rules/troll_warm.smk)

* Script: [`troll_warm.R`](https://github.com/sylvainschmitt/trollExp/blob/soil/scripts/troll_warm.R)

Run a TROLL warm up simulation (e.g. creation of a 600-years old mature forest).

## TROLL outputs

### [gather_warm](https://github.com/sylvainschmitt/trollExp/blob/main/rules/gather_warm.smk)

- Script:
  [`gather_warm.R`](https://github.com/sylvainschmitt/trollExp/blob/main/scripts/gather_warm.R)

Gather outputs of all experiments.

# Singularity

The whole workflow currently rely on the [`singularity-troll` image](https://github.com/sylvainschmitt/singularity-troll).

> need to add tidyterra, ggtern

# Data

## Soil

#### **SoilGrids**

250x250m soil information for the globe with quantified spatial uncertainty ([Poggio et al. 2021](https://soil.copernicus.org/articles/7/217/2021/)).
Available [online](https://files.isric.org/soilgrids/latest/data/) with an example [R script](https://git.wur.nl/isric/soilgrids/soilgrids.notebooks/-/blob/master/markdown/webdav_from_R.md).
We will use following variables for TROLL:

* silt: proportion of silt
* clay: proportion of clay
* sand: proportion of sand
* soc: soil organic content
* bdod: dry bulk density
* phh20: hydrogen ion activity in water  
* cec: cation exchange capacity at ph 7
