---
title: "trollExp - Run a TROLL experiment with phenology"
author: Sylvain Schmitt -
date: today
date-format: medium
format: 
  gfm:
    toc: true
    toc-depth: 2
---

```{r setup, include=FALSE}
rm(list = ls()) ; invisible(gc()) ; set.seed(42)
library(knitr)
library(tidyverse)
theme_set(bayesplot::theme_default())
opts_chunk$set(echo = F, message = F, warning = F,
               fig.height = 6, fig.width = 8,
               cache = F, cache.lazy = F)
```

[`singularity` & `snakemake`](https://github.com/sylvainschmitt/snakemake_singularity) 
workflow to run a TROLL experiment with phenology.

```{r dag}
#| fig.cap: "Workflow."
knitr::include_graphics("dag/dag.svg")
```

# Installation

-   [x] Python ≥3.5
-   [x] Snakemake ≥5.24.1
-   [x] Singularity ≥3.7.3

Once installed simply clone the workflow and go to the corresponding branch:

```{bash install}
#| eval: false
#| echo: true
git clone git@github.com:sylvainschmitt/trollExp.git
cd trollExp
git checkout pheno-fluxnet
```

# Usage

### Locally

```{bash usage}
#| eval: false
#| echo: true
snakemake --dag | dot -Tsvg > dag/dag.svg # dag locally
snakemake -j 1 --use-singularity --singularity-args # run locally
module load bioinfo/Snakemake/7.20.0 # for test on genobioinfo node
snakemake -np # dry run
sbatch job_genobioinfo.sh # run
```

# Workflow

## Data

### [rename](https://github.com/sylvainschmitt/trollExp/blob/pheno-fluxnet/rules/rename.py)

Rename and copy FLUXNET raw files.

### [format](https://github.com/sylvainschmitt/trollExp/blob/pheno-fluxnet/rules/format.py)

-   Script: [`format.R`](https://github.com/sylvainschmitt/trollExp/blob/pheno-fluxnet/scripts/format.R)

Format FLUXNET raw, species and soil data into TROLL inputs.

## Spin-up

### [prepare_spinup](https://github.com/sylvainschmitt/trollExp/blob/pheno-fluxnet/rules/prepare_spinup.py)

-   Script: [`prepare_spinup.R`](https://github.com/sylvainschmitt/trollExp/blob/pheno-fluxnet/scripts/prepare_spinup.R)

Prepare climate data as a TROLL input for a 600-years spin-up.

### [spinup](https://github.com/sylvainschmitt/trollExp/blob/pheno-fluxnet/rules/spinup.py)

-   Script: [`spinup.R`](https://github.com/sylvainschmitt/trollExp/blob/pheno-fluxnet/scripts/spinup.R)

Run a TROLL simulation for a 600-years spin-up.

## Run

### [prepare_run](https://github.com/sylvainschmitt/trollExp/blob/pheno-fluxnet/rules/prepare_run.py)

-   Script: [`prepare_run.R`](https://github.com/sylvainschmitt/trollExp/blob/pheno-fluxnet/scripts/prepare_run.R)

Prepare climate data as a TROLL for the run.

### [run](https://github.com/sylvainschmitt/trollExp/blob/pheno-fluxnet/rules/run.py)

-   Script: [`run.R`](https://github.com/sylvainschmitt/trollExp/blob/pheno-fluxnet/scripts/run.R)

Run a TROLL simulation.

### [run2](https://github.com/sylvainschmitt/trollExp/blob/pheno-fluxnet/rules/run2.py)

-   Script: [`run2.R`](https://github.com/sylvainschmitt/trollExp/blob/pheno-fluxnet/scripts/run2.R)

Run a second TROLL simulation for calibration.

# Singularity

The whole workflow currently rely on the [`singularity-troll` image](https://github.com/sylvainschmitt/singularity-troll).

# Data

-   [FLUXNET 2015](https://fluxnet.org/data/fluxnet2015-dataset/)
-   [Soil gathered by Isabelle](https://main--altpages.netlify.app/pheno#soils)
-   [Imputed species traits](https://main--altpages.netlify.app/biod-imp)
